---
title       : Simple BudgetFood Calculator
subtitle    : Predicting the percentage of household expenses used for food.
author      : JDScience
job         : Developing Data Products (February 2015)
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # tomorrow
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction to "BudgetFood Calculator"

Agenda
- What the Application Does
- How the Application Works
- Challenges Faced During Creation

--- .class #id

## What the Application Does

- Using the 'BudgetFood' dataset from the 'Ecdat' package
- Build a linear model to predict "wfood" (percentage of expenses spent on food)
- Allow application user to input their own values
- Allow user to "predict" their own "wfood" percentage

--- .class #id

## How the Application Works

Application consists of three files: ui.R, server.R, and helper.R
In the helper.R file, the data was loaded, and the model was built:
```{r, echo=FALSE}
library(Ecdat)
data(BudgetFood)
```
```{r}
model <- lm(wfood ~ totexp + age + size + town + sex, data = BudgetFood)
```

The model demonstrated good fit:

--- .class #id

```{r}
summary(model)
```

--- .class #id

## User Data Selection and Interaction

- As part of the application, several UI widgets were provided allowing the user
to input a set of values for the same variables built into the model. These controls
are displayed down the left column of the page.
- Changes to the widgets selected new values, which are displayed immediately to
the user in the main panel of the page.
- Under the row where the values are displayed is a button. This button is used
to run the 'predict' command, returning the predicted percentage of food spend
as a percentage.
- All the UI elements were defined in the ui.R file.

--- .class #id

## Calculation and Prediction

- The remainder of the application functionality, including calculations (specifically)
the 'predict' function were performed within the server.R file.
- The 'predict' command was initiated by the action button defined on the UI.
- The associated code looked like this:

```{r, eval=FALSE}
output$Xprediction <- renderPrint({
    input$predict
    isolate(percent(predict(model, newdata = data.frame(totexp = input$IscaledExp*200,
                                         age = input$Iage,
                                         size = input$Isize, 
                                         town = input$Itown, 
                                         sex = input$Igender,
                                         stringsAsFactors = TRUE))))
    })
```

--- .class #id

## Challenges and 'Solutions'

Making the currency relevant to our cultural environment
- The range of 'totexp' (total household expenses) was huge (~14,000 - ~11,000,000)
- The 'solution' was to scope the UI widget to a semi-reasonable range (10-10,000)
and the multiply the response by 200 before predicting the value.
- This 'hack' allowed the model to produce consistent values in a currency users
of the application could understand. The data was collected in Spain in 1980, so
the relative currency at the time was dramatically different.

--- .class #id 

## Challenge Part 2

Negative Predicted Values
- A direct scaling of the 'totexp' field proved tricky, resulting in predictions
of a negative percentage of expenses. The data ended up not being perfectly linear.
This can be seen to some extent in the plot on the following slide.
- The 'hack' described on the previous slide help normalize the currency values.
- The multiplier set to '200' helped mask the potential negative values from the
application.

--- .class #id

```{r}
plot(BudgetFood$wfood, BudgetFood$totexp,
     xlab = "% Expenses Spent on Food", ylab = "Total Household Expenses")
```

--- .class #id

## Conclusion

Thank you for viewing my presentation
- This presentation:
    - Described my Project
    - Explained how it worked
    - Discussed a couple challenges and how they were dealt with




