---
title       : Simple BudgetFood Calculator
subtitle    : Predicting the percentage of household expenses used for food.
author      : JDScience
job         : Developing Data Products (February 2015)
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # tomorrow
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction to "BudgetFood Calculator"

Functionality Included
- Using the 'BudgetFood' dataset from the 'Ecdat' package
- Build a linear model to predict "wfood" (percentage of expenses spent on food)
- Allow application user to input their own values
- "Predict" users own "wfood" percentage

--- .class #id

## How the Application Works

Application consists of three files: ui.R, server.R, and helper.R
In the helper.R file, the data was loaded, and the model was built:
```{r, echo=FALSE}
library(Ecdat)
data(BudgetFood)
```
```{r}
model <- lm(wfood ~ totexp + age + size + town + sex, data = BudgetFood)
```
The model demonstrated good fit (even though this page did not...):
```{r}
summary(model)
```

--- .class #id

## Calculation and Prediction

- The remainder of the application functionality, including calculations (specifically)
the 'predict' function were performed within the server.R file.
- The 'predict' command was initiated by the action button defined on the UI.
- The associated code looked like this:
```{r, eval=FALSE}
output$Xprediction <- renderPrint({
    input$predict
    isolate(percent(predict(model, newdata = data.frame(totexp = input$IscaledExp*200,
                                         age = input$Iage, size = input$Isize, 
                                         town = input$Itown, sex = input$Igender,
                                         stringsAsFactors = TRUE))))
    })
```
Note: The overall relationship between food cost and total expenses can be seen as follows...

--- .class #id

```{r}
plot(BudgetFood$wfood, BudgetFood$totexp,
     xlab = "% Expenses Spent on Food", ylab = "Total Household Expenses")
```
